---
title: "Mode of transportation prediction"
output: html_notebook
---
Import the dataset
```{r}
car_file <- read.csv("Cars.csv", header = T) # Import the dataset and store it as a dataframe
head(car_file) # Check the top few rows of our dataset to confirm the dataset is correctly loaded
str(car_file) # Check the structure of the dataset

# Converting the interger datatypes to factor as 3 variables are categorical with binary classification
car_file$Engineer <- as.factor(car_file$Engineer)
car_file$MBA <- as.factor(car_file$MBA)
car_file$license <- as.factor(car_file$license)

str(car_file) # Re-check the structure of the dataset after converting 3 variables
summary(car_file)
```
Identify missing values in the dataset
```{r}
library(VIM)
library(mice)
md.pattern(car_file)
mice_plot <- aggr(car_file, col=c('Aquamarine','Darkorchid'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(car_file), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))
```
Imputing missing value with MICE
```{r}
imputed_car_file <- mice(car_file, m=5, maxit = 5, method = 'logreg', seed = 50)
summary(imputed_car_file)
imputed_car_file$imp$MBA # Check what value was imputed after performing logistic regression on missing cell

imputed_car_file <- complete(imputed_car_file) # Create a dataframe from the model
md.pattern(imputed_car_file) # Re-check for any missing values in the dataset
```
Check for imbalance in dependent variable
```{r}
table(car_file$Transport)
(83/418)*100
(35/418)*100
(300/418)*100
```
Convert the dependent variable to a binary classification with 2 levels as we are to deal only with car.
```{r}
imputed_car_file$Transport <- ifelse(imputed_car_file$Transport == "2Wheeler", "0",
                                     ifelse(imputed_car_file$Transport == "Public Transport", "0",
                                            ifelse(imputed_car_file$Transport == "Car", "1", NA)))
```
Re-check the balance between the depedent variable
```{r}
table(imputed_car_file$Transport)
(383/418)*100
(35/418)*100
```
EDA - Exploratory data analysis
```{r}
imputed_car_file$Transport <- as.factor(imputed_car_file$Transport) # Convert the datatyoe of dependent variable to factor
str(imputed_car_file) # Check the structure of the dataset after conversion
```
EDA - continious variable plots - Histograms
```{r}
library(ggplot2)
ggplot(imputed_car_file, aes(x=Age)) + geom_histogram(bins = 80, binwidth = 0.5, color = "Darkorchid", fill= "Aquamarine")
ggplot(imputed_car_file, aes(x=Work.Exp)) + geom_histogram(bins = 80, binwidth = 0.5, color = "Darkorchid", fill= "Aquamarine")
ggplot(imputed_car_file, aes(x=Salary)) + geom_histogram(bins = 80, binwidth = 1, color = "Darkorchid", fill= "Aquamarine")
ggplot(imputed_car_file, aes(x=Distance)) + geom_histogram(bins = 80, binwidth = 0.5, color = "Darkorchid", fill= "Aquamarine")
```
EDA - continious variable plots - Boxplots
```{r}
attach(imputed_car_file)

boxplot(Age~Transport)

ggplot(imputed_car_file, aes(x = Age)) + 
  geom_density(aes(fill = Transport), alpha = 0.3) + 
  scale_color_manual(values = c("#868686FF", "#EFC000FF")) + 
  scale_fill_manual(values = c("darkturquoise", "lightcoral")) + xlim(-30,250)

boxplot(Work.Exp~Transport)

ggplot(imputed_car_file, aes(x = Work.Exp)) + 
  geom_density(aes(fill = Transport), alpha = 0.3) + 
  scale_color_manual(values = c("#868686FF", "#EFC000FF")) + 
  scale_fill_manual(values = c("darkturquoise", "lightcoral")) + xlim(-30,250)

boxplot(Salary~Transport)

ggplot(imputed_car_file, aes(x = Salary)) + 
  geom_density(aes(fill = Transport), alpha = 0.3) + 
  scale_color_manual(values = c("#868686FF", "#EFC000FF")) + 
  scale_fill_manual(values = c("darkturquoise", "lightcoral")) + xlim(-30,250)

boxplot(Distance~Transport)

ggplot(imputed_car_file, aes(x = Distance)) + 
  geom_density(aes(fill = Transport), alpha = 0.3) + 
  scale_color_manual(values = c("#868686FF", "#EFC000FF")) + 
  scale_fill_manual(values = c("darkturquoise", "lightcoral")) + xlim(-30,250)
```
EDA - continious varaible plots - Density plots
```{r}
ggplot(imputed_car_file, aes(x=Age)) + geom_density(color = "Orange") + geom_vline(aes(xintercept=mean(Age)),
            color="Purple", linetype="dashed", size=1)
ggplot(imputed_car_file, aes(x=Work.Exp)) + geom_density(color = "Orange") + geom_vline(aes(xintercept=mean(Work.Exp)),
            color="Purple", linetype="dashed", size=1)
ggplot(imputed_car_file, aes(x=Salary)) + geom_density(color = "Orange") + geom_vline(aes(xintercept=mean(Salary)),
            color="Purple", linetype="dashed", size=1)
ggplot(imputed_car_file, aes(x=Distance)) + geom_density(color = "Orange") + geom_vline(aes(xintercept=mean(Distance)),
            color="Purple", linetype="dashed", size=1)
```
EDA - continious variable plots - Histogram-Density plots
```{r}
ggplot(imputed_car_file, aes(x=Age)) +
 geom_histogram(aes(y=..density..), bins = 80, binwidth = 0.5, color = "Deepskyblue2", fill= "Aquamarine", alpha = 0.3) +
 geom_density(color = "Darkmagenta", alpha=0.7) + geom_vline(aes(xintercept=mean(Age)),
            color="Darkmagenta", linetype="dashed", size=1)

ggplot(imputed_car_file, aes(x=Work.Exp)) +
 geom_histogram(aes(y=..density..), bins = 80, binwidth = 0.5, color = "Deepskyblue2", fill= "Aquamarine", alpha = 0.3) +
 geom_density(color = "Darkmagenta", alpha=0.7) + geom_vline(aes(xintercept=mean(Work.Exp)),
            color="Darkmagenta", linetype="dashed", size=1)

ggplot(imputed_car_file, aes(x=Salary)) +
 geom_histogram(aes(y=..density..), bins = 80, binwidth = 1, color = "Deepskyblue2", fill= "Aquamarine", alpha = 0.3) +
 geom_density(color = "Darkmagenta", alpha=0.7) + geom_vline(aes(xintercept=mean(Salary)),
            color="Darkmagenta", linetype="dashed", size=1)

ggplot(imputed_car_file, aes(x=Distance)) +
 geom_histogram(aes(y=..density..), bins = 80, binwidth = 0.5, color = "Deepskyblue2", fill= "Aquamarine", alpha = 0.3) +
 geom_density(color = "Darkmagenta", alpha=0.7) + geom_vline(aes(xintercept=mean(Distance)),
            color="Darkmagenta", linetype="dashed", size=1)
```
EDA - continious variable plots - Jitter-Boxplots
```{r}
library(dplyr)
# Create separate boxplots for each attribute
imputed_car_file %>%
 ggplot(aes(x="", y=Age)) +
 geom_boxplot(varwidth=T, outlier.colour = "Deepskyblue2", fill="Darkorchid4") + coord_flip() +  geom_jitter(colour="Darkorange",width=0.2,alpha=0.2)

imputed_car_file %>%
 ggplot(aes(x="", y=Work.Exp)) +
 geom_boxplot(varwidth=T, outlier.colour = "Deepskyblue2", fill="Darkorchid4") + coord_flip() +  geom_jitter(colour="Darkorange",width=0.2,alpha=0.2)

imputed_car_file %>%
 ggplot(aes(x="", y=Salary)) +
 geom_boxplot(varwidth=T, outlier.colour = "Deepskyblue2", fill="Darkorchid4") + coord_flip() +  geom_jitter(colour="Darkorange",width=0.2,alpha=0.2)

imputed_car_file %>%
 ggplot(aes(x="", y=Distance)) +
 geom_boxplot(varwidth=T, outlier.colour = "Deepskyblue2", fill="Darkorchid4") + coord_flip() +  geom_jitter(colour="Darkorange",width=0.2,alpha=0.2)
```
EDA - categorical variable plotting
```{r}
ggplot(imputed_car_file, aes(x = Gender, fill = Transport)) + 
  geom_bar(width = 0.25, alpha=0.5) + 
  scale_fill_manual(values = c('darkturquoise', 'lightcoral'))

ggplot(imputed_car_file, aes(x = Engineer, fill = Transport)) + 
  geom_bar(width = 0.25, alpha=0.5) + 
  scale_fill_manual(values = c('darkturquoise', 'lightcoral'))

ggplot(imputed_car_file, aes(x = MBA, fill = Transport)) + 
  geom_bar(width = 0.25, alpha=0.5) + 
  scale_fill_manual(values = c('darkturquoise', 'lightcoral'))

ggplot(imputed_car_file, aes(x = license, fill = Transport)) + 
  geom_bar(width = 0.25, alpha=0.5) + 
  scale_fill_manual(values = c('darkturquoise', 'lightcoral'))
```
Table distribution of categorical variables
```{r}
prop.table(table(Gender,Transport),1)*100
prop.table(table(Engineer,Transport),1)*100
prop.table(table(MBA,Transport),1)*100
prop.table(table(license,Transport),1)*100
```
Preparing data for analysis - Splitting the dataset
```{r}
library(DMwR)
library(caret)
library(caTools) # import library caTools for splitting data
set.seed(48) # Set the seed to repeat the same sample set for each iteration
car_split <- sample.split(imputed_car_file$Transport, SplitRatio = 0.7) # Split the dataset into 70-30
car_train <- subset(imputed_car_file, car_split==TRUE) # Create subset of train data from dataset
car_test <- subset(imputed_car_file, car_split==FALSE) # Create subset of test data from dataset
prop.table(table(car_train$Transport)) # Check the distribution of the dependent variable in train data
prop.table(table(car_test$Transport)) # Check the distribution of the dependent variable in test data
```
Preparing the data for analysis - SMOTE
```{r}
table(car_train$Transport)
# Event rate of minority class is 24/292 which is 8.2%
balanced_car <- SMOTE(Transport~., data = car_train, perc.over = 1500, k=9, perc.under = 300)
table(balanced_car$Transport)
# Event rate of the minority class is 384/1464 which is 46.4%
```
Splitting the SMOTE dataset into train and test
```{r}
set.seed(100) # Set the seed to repeat the same sample set for each iteration
smote_car_split <- sample.split(balanced_car$Transport, SplitRatio = 0.7) # Split the dataset into 70-30
smote_car_train <- subset(balanced_car, smote_car_split==TRUE) # Create subset of train data from dataset
smote_car_test <- subset(balanced_car, smote_car_split==FALSE) # Create subset of test data from dataset
table(smote_car_train$Transport)
```
Logistic Regression
```{r}
LR_model1 = glm(Transport ~ ., data = smote_car_train, 
                   family = binomial(link="logit"))

summary(LR_model1)
```
Check for multi-collinearity
```{r}
library(car)
vif(LR_model1)
```
Refining the logistic regression model
```{r}
LR_model2 = glm(Transport ~ . -Age -Work.Exp, data = smote_car_train, 
                   family = binomial(link="logit"))

summary(LR_model2)
```
Re-check multi-collinearity with refined model
```{r}
vif(LR_model2)
```
Model significance tests
```{r}
library(lmtest)
library(pscl)

# Likelihood ratio test
lrtest(LR_model2)

# Pseudo R-squared
pR2(LR_model2)

#Odds Ratio
exp(coef(LR_model2))

#Probability
exp(coef(LR_model2))/(1+exp(coef(LR_model2)))
```
Model performance metrics - Train data
```{r}
pred_train_LR <- predict(LR_model2, data=smote_car_train, type = "response")
y_pred_num = ifelse(pred_train_LR>0.5,1,0)
y_pred = factor(y_pred_num, levels=c(0,1))
y_actual = smote_car_train$Transport
confusionMatrix(y_pred,y_actual,positive="1")
```
Model performance metrics 2 - Train data
```{r}
library(ROCR)
smote_car_train.roc <- prediction(pred_train_LR, smote_car_train$Transport)
plot(performance(smote_car_train.roc, "tpr", "fpr"), 
     col = "red", main = "ROC Curve for train data")
abline(0, 1, lty = 8, col = "blue")

# AUC
smote_car_train.auc = performance(smote_car_train.roc, "auc")
smote_car_train.area = as.numeric(slot(smote_car_train.auc, "y.values"))
smote_car_train.area

# KS
smote_car_ks.train <- performance(smote_car_train.roc, "tpr", "fpr")
smote_car_train.ks <- max(attr(smote_car_ks.train, "y.values")[[1]] - (attr(smote_car_ks.train, "x.values")[[1]]))
smote_car_train.ks

# Gini
smote_car_train.gini = (2 * smote_car_train.area) - 1
smote_car_train.gini
```
Model performance metrics - Test data
```{r}
pred_test_LR = predict(LR_model2, newdata=smote_car_test, type="response")
y_pred_num = ifelse(pred_test_LR>0.35,1,0)
y_pred = factor(y_pred_num, levels=c(0,1))
y_actual = smote_car_test$Transport
confusionMatrix(y_pred,y_actual,positive="1")
```

```{r}
library(ROCR)
smote_car_test.roc <- prediction(pred_test_LR, smote_car_test$Transport)
plot(performance(smote_car_test.roc, "tpr", "fpr"), 
     col = "red", main = "ROC Curve for test data")
abline(0, 1, lty = 8, col = "blue")

# AUC
smote_car_test.auc = performance(smote_car_test.roc, "auc")
smote_car_test.area = as.numeric(slot(smote_car_test.auc, "y.values"))
smote_car_test.area

# KS
smote_car_ks.test <- performance(smote_car_test.roc, "tpr", "fpr")
smote_car_test.ks <- max(attr(smote_car_ks.test, "y.values")[[1]] - (attr(smote_car_ks.test, "x.values")[[1]]))
smote_car_test.ks

# Gini
smote_car_test.gini = (2 * smote_car_test.area) - 1
smote_car_test.gini
```

AUC and KS plots
```{r}
library(InformationValue)
plotROC(actuals = smote_car_test$Transport, predictedScores = pred_test_LR)
pred_test_LR

ks_stat <- InformationValue::ks_stat(actuals = smote_car_test$Transport, predictedScores = pred_test_LR)
ks_stat
ks_plot(actuals = smote_car_test$Transport, predictedScores = pred_test_LR)
```
KNN - K nearest neighbour
```{r}
#Convert all the categorical variables to boolean

scale = preProcess(smote_car_train, method = "range")

train.norm.data = predict(scale, smote_car_train)
test.norm.data = predict(scale, smote_car_test)

knn_fit = train(Transport ~., data = train.norm.data, method = "knn",
                 trControl = trainControl(method = "cv", number = 3),
                 tuneLength = 10)

knn_fit$bestTune$k
plot(knn_fit)
```
Performance metrics - Test data
```{r}
library(class)
knn_fit<- knn(train = train.norm.data[,-c(2,3,4,8)], test = test.norm.data[,-c(2,3,4,8)], cl= train.norm.data[,9],k = 5,prob=TRUE) 
#dim(train.norm.data)
#dim(test.norm.data)
table(test.norm.data[,9],knn_fit)
```
NB - Naive Bayes model building
```{r}
library(e1071)
NB = naiveBayes(x=train.norm.data[,1:8], y=as.factor(train.norm.data[,9]))
# Performance metrics (out-of-the-sample)
pred_nb<-predict(NB,newdata = test.norm.data[,1:8])

table(test.norm.data[,9],pred_nb)
```
Bagging
```{r}
library(gbm)         # basic implementation using AdaBoost
library(xgboost)      # a faster implementation of a gbm
library(ipred)
library(rpart)

car_bagging <- bagging(Transport ~.,
                           data=smote_car_train,
                           control=rpart.control(maxdepth=5, minsplit=4))


smote_car_test$pred.class <- predict(car_bagging, smote_car_test)


#gd_test$pred.class<- ifelse(gd_test$pred.class<0.5,0,1)


#confusionMatrix(data=factor(gd_test$pred.class),
#                reference=factor(gd_test$Class),
#                positive='1')

table(smote_car_test$Transport,smote_car_test$pred.class)
```

```{r}
## work around bug in gbm 2.1.1
predict.gbm <- function (object, newdata, n.trees, type = "link", single.tree = FALSE, ...) {
  if (missing(n.trees)) {
    if (object$train.fraction < 1) {
      n.trees <- gbm.perf(object, method = "test", plot.it = FALSE)
    }
    else if (!is.null(object$cv.error)) {
      n.trees <- gbm.perf(object, method = "cv", plot.it = FALSE)
    }
    else {
      n.trees <- length(object$train.error)
    }
    cat(paste("Using", n.trees, "trees...\n"))
    gbm::predict.gbm(object, newdata, n.trees, type, single.tree, ...)
  }
}
```

Boosting
```{r}
car_boosting <- gbm(
  formula = Transport ~ .,
  distribution = "bernoulli",
  data = smote_car_train,
  n.trees = 100,
  interaction.depth = 1,
  shrinkage = 0.001,
  n.cores = NULL, # will use all cores by default
  verbose = FALSE
  )  

smote_car_test$pred.class <- predict(car_boosting, smote_car_test, type = "response")

table(smote_car_test$Transport,smote_car_test$pred.class)
```


