---
title: "Time Series Forecasting Project"
output: html_notebook
---
Importing the required packages
```{r}
library(forecast) # Importing the dataset present in the Forecast package
library(timeSeries) # Importing the essential functions used for time series data manipulation
View(gas) # View the time series dataset, with annual monthly gas production in Australia
ts.plot(gas, col = "DarkOrchid") # Plotting the time series to observe the components
```
Storing the dataset as a time series object
```{r}
# As the dataset is a monthly time series, the frequency is mentioned as 12.
gas_ts <- ts(gas, start = c(1956,1), frequency = 12) # Saving the dataset as a time series object
# The Start date for the object is the beginning of the dataset and no end date is required as entire dataset is considered for forecasting.
```
Checking periodicity
```{r}
library(xts) # Package which holds the periodicity function
periodicity(gas_ts) # Check periodicity of the time series object
```

Exploring the time series object
```{r}
start(gas_ts) # Returns the start year and month of the time series object
end(gas_ts) # Returns the end year and month of the time series object
frequency(gas_ts) # Returns the frequenct of the time series object
summary(gas_ts) # Returns the statistical summary of the time series object
cycle(gas_ts) # Returns the cycle of the time series object

gas_ts_Q <- aggregate(gas_ts, nfrequency = 4) # Aggregating time series object to quarterly
gas_ts_Y <- aggregate(gas_ts, nfrequency = 1) # Aggregating time series object to yearly
```
Check for missing values
```{r}
library(imputeTS)
 plotNA.distribution(gas_ts, colPoints = "steelblue",
  colBackgroundMV = "indianred2", main = "Distribution of NAs",
  xlab = "Time", ylab = "Value", pch = 20, cexPoints = 0.8,
  col = "black")
```
Plotting the time series objects
```{r}
plot.ts(gas_ts, main = "Australian monthly gas production 1956-1995", xlab = "Year", ylab = "Units") # Monthly time series plot
plot.ts(gas_ts_Q, main = "Australian quarterly gas production 1956-1995", xlab = "Year", ylab = "Units") # Quarterly time series plot
plot.ts(gas_ts_Y, main = "Australian yearly gas production 1956-1995", xlab = "Year", ylab = "Units") # Yearly time series plot

# Advanced plots
ggseasonplot(gas_ts, year.labels.left = T, year.labels = T, col = 1:40, main = "Australian seasonal gas production", xlab = "Months", ylab = "Units") # Seasonal time series plot
monthplot(gas_ts, main = "Australian gas production - month wise 1956-1995", xlab = "Months", ylab = "Units") # Monthly time series plot
boxplot(gas_ts ~ cycle(gas_ts), col = c("DarkOrange", "Purple"), main = "Australian monthly gas production - Boxplot", xlab = "Months", ylab = "Units", horizontal = T)
```
Decomposition of the time series
```{r}
decomp_gas_add <- decompose(gas_ts, type = "additive") # Decomposing the additive model - S + T + N
plot(decomp_gas_add) # Decomposition plot of the additive model
decomp_gas_mul <- decompose(gas_ts, type = "multiplicative") # Decomposing the multiplicative model - S * T * N
plot(decomp_gas_mul) # Decomposition plot of the multiplicative model
# decomp_gas_log <- log(gas_ts)
# plot.ts(decomp_gas_log)
```
Individual component plots
```{r}
plot(decomp_gas_add$trend) # Trend plot for additive model of time series
plot(decomp_gas_add$seasonal) # Seasonal plot for additive model of time series
plot(decomp_gas_add$random) # Error/Random plot for additive model of time series

plot(decomp_gas_mul$trend) # Trend plot for multiplicative model of time series
plot(decomp_gas_mul$seasonal) # Seasonal plot for multiplicative model of time series
plot(decomp_gas_mul$random) # Error/Random plot for multiplicative model of time series
```
Advanced decomposition of time series using STL
```{r}
plot(stl(gas_ts, s.window = "periodic")) # Decomposing the components of the time series - Trend, Seasonality and Noise
decomp_gas_stl <- stl(gas_ts, s.window = 7) # Setting seasonal window to 3
plot(decomp_gas_stl)
deseason_gas <- decomp_gas_stl$time.series[,2]+decomp_gas_stl$time.series[,3]
ts.plot(deseason_gas, gas_ts, col=c("DarkOrchid", "Orange"), main="Comparison of Gas produced and Deseasoned Series")
```
Differentiating trend and seasonality from time series data
```{r}
library(ggfortify)
decomp_gas_add$seasonal
decomp_gas_mul$seasonal
decomp_gas_stl
```
Seasonal and Trend analysis for Additive model
```{r}
autoplot(decomp_gas_add$trend)
autoplot(decomp_gas_add$seasonal)
autoplot(decomp_gas_add)
```
Seasonal and Trend analysis for Mutiplicative model
```{r}
autoplot(decomp_gas_mul$trend)
autoplot(decomp_gas_mul$seasonal)
autoplot(decomp_gas_mul)
```
Extracting residuals from additive model
```{r}
res_add <- decomp_gas_add$x - decomp_gas_add$seasonal - decomp_gas_add$trend
res_add
```
Extracting residuals from additive model
```{r}
res_mul <- decomp_gas_mul$x - decomp_gas_mul$seasonal - decomp_gas_mul$trend
res_mul
```
Residual plots
```{r}
qqnorm(res_add)
qqnorm(res_mul)
autoplot(res_add)
autoplot(res_mul)
```
Augmented Dickey-Fuller test for stationarity
```{r}
library(tseries)
adf.test(gas_ts)
```
Kwiatkowski–Phillips–Schmidt–Shin (KPSS) test
```{r}
kpss.test(gas_ts, null = "Trend")
```
Split the dataset into training and testing data
```{r}
gas_train <- window(gas_ts, end = c(1994,8))
gas_test <- window(gas_ts, start = c(1994, 9))

gas_train
gas_test
```
Holt's Model
```{r}
gas_holt <- holt(gas_train, type = "additive", h = 12)
autoplot(gas_holt)

gas_holt$model
accuracy(gas_holt, gas_test)
```
Holt's Winter Model
```{r}
library(forecast)
gas_HW <- HoltWinters(gas_train, gamma = FALSE)
gas_HW
plot(gas_HW)

gas_HWF <- forecast(gas_HW, h = 12)
plot(gas_HWF)
```
ARIMA modelling - Manual ARIMA
```{r}
acf(gas_ts)
pacf(gas_ts)
```
Differencing the time series
```{r}
gas_ds <- seasadj(decomp_gas_stl)
plot(gas_ds)
gas_d1 <- diff(gas_ds, differences = 1)
plot(gas_d1)
adf.test(gas_d1, alternative = "stationary")
```
ACF and PACF for stationary series
```{r}
acf(gas_d1, main = "ACF for monthly Australian gas production")
pacf(gas_d1)
```
Splitting again into train and test
```{r}
gas_d1_train <- window(gas_d1, end = c(1994,8))
gas_d1_test <- window(gas_d1, start = c(1994, 9))
```
ARIMA model - Auto ARIMA
```{r}
gas_ARIMA <- auto.arima(gas_d1_train, seasonal = FALSE)
gas_ARIMA
tsdisplay(residuals(gas_ARIMA))
```
Ljung-Box test
```{r}
Box.test(gas_ARIMA$residuals)
```
Forecasting with ARIMA
```{r}
fcast <- forecast(gas_ARIMA, h=12)
plot(fcast)
```
Accuracy
```{r}
accuracy(fcast, gas_d1_test)
```

