---
title: "Project 2 - Predicting loan defaults"
output: html_notebook
---
Problem Statement - A bank wants to implement a model which will predict the potential defaulters based on previous history.
Dataset - Bank loan data
Variables - 9 (8 independent & 1 dependent)
Dep. Variable - default (Categorical)
Observations - 850

Setting the working directory & loading the dataset
```{r}
setwd("~/OneDrive/EdWisor Assignments/Project - 2") # Setting the working directory
getwd() # Fetching the current working directory

train_loan <- read.csv("bank-loan.csv", header = TRUE) # Importing train dataset
str(train_loan) # Checking the structure of train dataset
summary(train_loan) # Checking the summary of train dataset

apply(train_loan, 2, function(x) length(table(x))) # Checking unique counts for each variable
```
Phase 1 - Data Preprocessing
```{r}
library(DataExplorer)
library(tidyr)
library(dplyr)

apply(train_loan, 2,  function(x) {sum(is.na(x))}) # Checking the missing values from each column
plot_missing(train_loan) # Plotting the missing values chart for each column

test_loan <- train_loan[is.na(train_loan$default),] # Created a separate test dataframe from missing observations.

train_loan <- drop_na(train_loan) # Deleting the missing observations from train dataset
```
Checking the structure of the dataset
```{r}
str(train_loan) # Checking the structure of train dataset

str(test_loan) # Checking the structure of test dataset
```
```{r}
train_loan$ed <- as.factor(train_loan$ed) # Converting the education variable to factor
train_loan$default <- as.factor(train_loan$default) # Converting the default variable to factor
```
Data Exploration
```{r}
# We'll prepare our data for visualisations by converting each variable to its appropriate datatype

train_cont <- train_loan[,c(1,3:8)] # Creating a separate dataframe for continuous variable analysis
```
Visualisations
```{r}
library(ggplot2)

# We'll plot univariate categorical variables

ggplot(train_loan, aes(x = ed, fill=ed)) + geom_bar(color = "azure4", bins = 100) + labs(title="Count of no. of applicants by education level", x = "Education level") + scale_fill_brewer(palette="Greens") + theme_minimal() # Applicant count by education

# We'll plot univariate continuous variables

ggplot(train_cont, aes(x = age)) + geom_histogram(fill = "skyblue", color = "azure4", bins = 100) + labs(title="Age distribution of applicants", x = "Age of the applicants") + scale_fill_brewer(palette="Greens") + theme_minimal() # Distribution of age

ggplot(train_cont, aes(x = employ)) + geom_histogram(fill = "skyblue", color = "azure4", bins = 100) + labs(title="No. of years of employment", x = "Employment in years") + scale_fill_brewer(palette="Greens") + theme_minimal() # Employment in years

ggplot(train_cont, aes(x = address)) + geom_histogram(fill = "skyblue", color = "azure4", bins = 100) + labs(title="Locality distribution of applicants", x = "Address of the applicants") + scale_fill_brewer(palette="Greens") + theme_minimal() # Distribution of address

ggplot(train_cont, aes(x = income)) + geom_histogram(fill = "skyblue", color = "azure4", bins = 100) + labs(title="Income distribution of applicants", x = "Income of the applicants") + scale_fill_brewer(palette="Greens") + theme_minimal() # Distribution of income

ggplot(train_cont, aes(x = debtinc)) + geom_histogram(fill = "skyblue", color = "azure4", bins = 100) + labs(title="Debt to income ratio of applicants", x = "DTI of the applicants") + scale_fill_brewer(palette="Greens") + theme_minimal() # DTI ratio

ggplot(train_cont, aes(x = creddebt)) + geom_histogram(fill = "skyblue", color = "azure4", bins = 100) + labs(title="Debt to credit ratio of applicants", x = "DTC of the applicants") + scale_fill_brewer(palette="Greens") + theme_minimal() # DTC ratio

ggplot(train_cont, aes(x = othdebt)) + geom_histogram(fill = "skyblue", color = "azure4", bins = 100) + labs(title="Other debt of applicants", x = "Other debts of the applicants") + scale_fill_brewer(palette="Greens") + theme_minimal() # Other debts
```
Boxplots
```{r}
library(viridis)

ggplot(train_cont, aes(x="", y=age))+ geom_boxplot(fill = "#45B8AC", outlier.color = "orange")+ 
  geom_jitter(aes(color = "azure4"), width = 0.15, height = 0.35) + coord_flip()+
  scale_color_viridis(discrete = T, option = "D", alpha = 0.6)+
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Box-plot of Age") +
  xlab("Spread") + ylab("Age")+ theme(
    plot.title = element_text(size=14, face="plain"),
    axis.title.x = element_text(size=14, face="plain"),
    axis.title.y = element_text(size=14, face="plain"))

ggplot(train_cont, aes(x="", y=income))+ geom_boxplot(fill = "#45B8AC", outlier.color = "orange")+ 
  geom_jitter(aes(color = "azure4"), width = 0.15, height = 0.35) + coord_flip()+
  scale_color_viridis(discrete = T, option = "D", alpha = 0.6)+
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Box-plot of Income") +
  xlab("Spread") + ylab("Income")+ theme(
    plot.title = element_text(size=14, face="plain"),
    axis.title.x = element_text(size=14, face="plain"),
    axis.title.y = element_text(size=14, face="plain"))

ggplot(train_cont, aes(x="", y=debtinc))+ geom_boxplot(fill = "#45B8AC", outlier.color = "orange")+ 
  geom_jitter(aes(color = "azure4"), width = 0.15, height = 0.35) + coord_flip()+
  scale_color_viridis(discrete = T, option = "D", alpha = 0.6)+
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Box-plot of Debt to income ratio") +
  xlab("Spread") + ylab("DTI Ratio")+ theme(
    plot.title = element_text(size=14, face="plain"),
    axis.title.x = element_text(size=14, face="plain"),
    axis.title.y = element_text(size=14, face="plain"))

ggplot(train_cont, aes(x="", y=creddebt))+ geom_boxplot(fill = "#45B8AC", outlier.color = "orange")+ 
  geom_jitter(aes(color = "azure4"), width = 0.15, height = 0.35) + coord_flip()+
  scale_color_viridis(discrete = T, option = "D", alpha = 0.6)+
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Box-plot of Debt to credit ratio") +
  xlab("Spread") + ylab("DTC Ratio")+ theme(
    plot.title = element_text(size=14, face="plain"),
    axis.title.x = element_text(size=14, face="plain"),
    axis.title.y = element_text(size=14, face="plain"))

ggplot(train_cont, aes(x="", y=othdebt))+ geom_boxplot(fill = "#45B8AC", outlier.color = "orange")+ 
  geom_jitter(aes(color = "azure4"), width = 0.15, height = 0.35) + coord_flip()+
  scale_color_viridis(discrete = T, option = "D", alpha = 0.6)+
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Box-plot of Other debts") +
  xlab("Spread") + ylab("Other debts")+ theme(
    plot.title = element_text(size=14, face="plain"),
    axis.title.x = element_text(size=14, face="plain"),
    axis.title.y = element_text(size=14, face="plain"))

```

Bi-variate Visualisations
```{r}
# We'll plot bivariate visualisations to better understand relationship between two variables

ggplot(train_loan, aes(x = age, y = income)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Age",
       y = "Applicant's Income", title = "Age vs Income")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = age, y = employ)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Age",
       y = "Applicant's Employment", title = "Age vs Employment")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = age, y = debtinc)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Age",
       y = "Applicant's DTI", title = "Age vs DTI ratio")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = age, y = creddebt)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Age",
       y = "Applicant's DTC", title = "Age vs DTC ratio")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = age, y = othdebt)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Age",
       y = "Applicant's Debts", title = "Age vs Other Debt")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = income, y = employ)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Income",
       y = "Applicant's Employment", title = "Income vs Employment")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = income, y = debtinc)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Income",
       y = "Applicant's DTI", title = "Income vs DTI Ratio")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = income, y = creddebt)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Income",
       y = "Applicant's DTC", title = "Income vs DTC Ratio")+ geom_smooth(method = "lm", colour = "darkmagenta")

ggplot(train_loan, aes(x = income, y = othdebt)) + geom_point(color="turquoise", size = 2, alpha=.8) +  labs(x = "Income",
       y = "Applicant's Other Debts", title = "Income vs Other Debts")+ geom_smooth(method = "lm", colour = "darkmagenta")

```
Correlation
```{r}
library(corrplot)
library(PerformanceAnalytics)
library(moments)

cor_mat <- cor(train_loan[,c(1,3:8)], method = c("pearson", "kendall", "spearman"), use = "complete.obs")
corrplot(cor_mat, method = "number")
chart.Correlation(train_loan[,c(1,3:8)], histogram=TRUE, pch=19)
```
Normality check
```{r}
library(igraph)
library(car)
library(MASS)

qqPlot(train_loan$age) # Plotting QQ plot for age

truehist(train_loan$age) # Plotting the histogram to check the tails
lines(density(train_loan$age))

dt=density(train_loan$age)
plot(dt,main="Distribution of Age") # Plotting density plot for age
polygon(dt,col="orange",border="darkmagenta")

#-------------
qqPlot(train_loan$employ) # Plotting QQ plot for employment

truehist(train_loan$employ) # Plotting the histogram to check the tails
lines(density(train_loan$employ))

dt=density(train_loan$employ)
plot(dt,main="Distribution of employment") # Plotting density plot for employment
polygon(dt,col="orange",border="darkmagenta")

#-------------
qqPlot(train_loan$income) # Plotting QQ plot for income

truehist(train_loan$income) # Plotting the histogram to check the tails
lines(density(train_loan$income))

dt=density(train_loan$income)
plot(dt,main="Distribution of income") # Plotting density plot for income
polygon(dt,col="orange",border="darkmagenta")

#-------------
qqPlot(train_loan$debtinc) # Plotting QQ plot for DTI Ratio

truehist(train_loan$debtinc) # Plotting the histogram to check the tails
lines(density(train_loan$debtinc))

dt=density(train_loan$debtinc)
plot(dt,main="Distribution of DTI ratio") # Plotting density plot for DTI Ratio
polygon(dt,col="orange",border="darkmagenta")

#-------------
qqPlot(train_loan$creddebt) # Plotting QQ plot for DTC Ratio

truehist(train_loan$creddebt) # Plotting the histogram to check the tails
lines(density(train_loan$creddebt))

dt=density(train_loan$creddebt)
plot(dt,main="Distribution of DTC ratio") # Plotting density plot for DTC Ratio
polygon(dt,col="orange",border="darkmagenta")

#-------------
qqPlot(train_loan$othdebt) # Plotting QQ plot for Other Debt

truehist(train_loan$othdebt) # Plotting the histogram to check the tails
lines(density(train_loan$othdebt))

dt=density(train_loan$othdebt)
plot(dt,main="Distribution of Other Debt") # Plotting density plot for Other Debt
polygon(dt,col="orange",border="darkmagenta")
```
Applying Log transformation
```{r}
train_loan$age=log1p(train_loan$age) # Applying log transformation on fare amount - Train
train_loan$employ=log1p(train_loan$employ) # Applying log transformation on distance - Train
train_loan$income=log1p(train_loan$income) # Applying log transformation on distance - Train
train_loan$address=log1p(train_loan$address) # Applying log transformation on distance - Train
train_loan$debtinc=log1p(train_loan$debtinc) # Applying log transformation on distance - Train
train_loan$creddebt=log1p(train_loan$creddebt) # Applying log transformation on distance - Train
train_loan$othdebt=log1p(train_loan$othdebt) # Applying log transformation on distance - Train

test_loan$age=log1p(test_loan$age) # Applying log transformation on fare amount - Test
test_loan$employ=log1p(test_loan$employ) # Applying log transformation on distance - Test
test_loan$income=log1p(test_loan$income) # Applying log transformation on distance - Test
test_loan$address=log1p(test_loan$address) # Applying log transformation on distance - Test
test_loan$debtinc=log1p(test_loan$debtinc) # Applying log transformation on distance - Test
test_loan$creddebt=log1p(test_loan$creddebt) # Applying log transformation on distance - Test
test_loan$othdebt=log1p(test_loan$othdebt) # Applying log transformation on distance - Test
```
Splitting the data
```{r}
library(caret)
library(caTools)

set.seed(100)
train_split <- createDataPartition(y=train_loan$default, p = 0.75, list = F)
train_data <- train_loan[train_split,]
test_data <- train_loan[-train_split,]
```
Data Balancing
```{r}
library(DMwR)

table(train_data$default) # Checking the distribution of classes in our dependent variable

train_data.sm <- SMOTE(default ~., train_data, perc.over = 150, k = 5, perc.under = 200) # Applying SMOTE to balance the classes

table(train_data.sm$default) # Checking the distribution of classes in our dependent variable after SMOTE
```
Logistic Regression
```{r}
logreg <- glm(default~., data=train_data.sm, family = binomial)
summary(logreg)
```
Prediction - LR
```{r}
#install.packages("Metrics")
library(Metrics)
library(ROCR)
library(InformationValue)

pred_test <- predict(logreg, test_data, type = 'response')

pred_LR <- ifelse(pred_test > 0.5, 1, 0)

pred_LR <- as.factor(pred_LR)

table(pred_LR, test_data$default)

#confusionMatrix(pred_LR, test_data$default)

roc <- prediction(pred_test, test_data$default)
as.numeric(performance(roc, "auc")@y.values)
auc <- performance(roc, "tpr", "fpr")
plot(auc)

optimalCutoff(actuals = test_data$default, predictedScores = as.numeric(pred_test)) # Checking the optimal cutoff 
plotROC(actuals = test_data$default, predictedScores = as.numeric(pred_test)) # Plotting the ROC curve
```
K-Nearest Neighbor
```{r}
library(class)

knn_1<-knn(train_data.sm,test_data,k=20,cl=train_data.sm$default)

knn_1
```
Prediction - KNN
```{r}
pred_knn <- predict(knn_1, test_data)

table(pred_knn, test_data$default)

#confusionMatrix(knn_1, test_data$default)

roc_1 <- prediction(knn_1, test_data$default)
as.numeric(performance(roc_1, "auc")@y.values)
auc_1 <- performance(roc_1, "tpr", "fpr")
plot(auc_1)

optimalCutoff(actuals = test_data$default, predictedScores = as.numeric(pred_knn)) # Checking the optimal cutoff 
plotROC(actuals = test_data$default, predictedScores = as.numeric(pred_knn)) # Plotting the ROC curve
```

Naive Bayes
```{r}
# Build Naive Bayes model with all variables
library(GGally)
library(e1071)
nbm = naiveBayes(default~., data = train_data.sm)

nbm
```
Prediction - NB
```{r}
pred_nb <- predict(nbm, test_data)

table(pred_nb, test_data$default)

roc_2 <- prediction(pred_nb, test_data$default)
as.numeric(performance(roc_2, "auc")@y.values)
auc_2 <- performance(roc_2, "tpr", "fpr")
plot(auc_2)

optimalCutoff(actuals = test_data$default, predictedScores = as.numeric(pred_nb)) # Checking the optimal cutoff 
plotROC(actuals = test_data$default, predictedScores = as.numeric(pred_nb)) # Plotting the ROC curve
```
CART - Classification Trees
```{r}
#install.packages("tree")
library(rpart)
library(rpart.plot)
library(rattle)
library(tree)

#tree_1 <- tree(default~., data = train_data.sm)
#summary(tree_1)

tree_2 <- rpart(default~., data = train_data.sm, method = "class")
tree_2
rpart.plot(tree_2)
```
Prediction - CART
```{r}
pred_cart <- predict(tree_2, test_data[-9], type = "class")
table(test_data$default, pred_cart)

roc_3 <- prediction(as.numeric(pred_cart), as.numeric(test_data$default))
as.numeric(performance(roc_3, "auc")@y.values)
auc_3 <- performance(roc_3, "tpr", "fpr")
plot(auc_3)

optimalCutoff(actuals = test_data$default, predictedScores = as.numeric(pred_cart)) # Checking the optimal cutoff 
plotROC(actuals = test_data$default, predictedScores = as.numeric(pred_cart)) # Plotting the ROC curve
```
Random Forest
```{r}
library(randomForest)
set.seed(100)
rf <- train(default ~ ., data = train_data.sm, method = 'rf', trControl = trainControl(method = 'cv', number = 10))
rf
```
Prediction - RF
```{r}
pred_rf <- predict(rf, newdata = test_data)
table(test_data$default, pred_rf)

#confusionMatrix(pred_rf, test_data$default)

roc_4 <- prediction(as.numeric(pred_rf), as.numeric(test_data$default))
as.numeric(performance(roc_4, "auc")@y.values)
auc_4 <- performance(roc_4, "tpr", "fpr")
plot(auc_4)

optimalCutoff(actuals = test_data$default, predictedScores = as.numeric(pred_rf)) # Checking the optimal cutoff 
plotROC(actuals = test_data$default, predictedScores = as.numeric(pred_rf)) # Plotting the ROC curve
```
